Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-04-26 doc/paper.Rmd
pics updated

errorsCV2 <- rbind(errorrate(cvcs2, 0.001), 
                errorrate(cvcs2, 0.005), 
                errorrate(cvcs2, 0.01), 
                errorrate(cvcs2, 0.05))
errorsCV2$method <- "CS2"

errorsCV1 <- rbind(errorrate(cvcs1, 0.001), 
                errorrate(cvcs1, 0.005), 
                errorrate(cvcs1, 0.01), 
                errorrate(cvcs1, 0.05))
errorsCV1$method <- "CS1"

errorsCV <- rbind(errorsCV1, errorsCV2)
errorsCV %>% 
  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
  geom_point(aes(shape=factor(wv)), size=2.5) +
  geom_smooth(aes(group=alpha), se=FALSE, method="loess", span=1,  size=.7) +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha),
                      palette="Set2") +
  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
  xlab(expression("Window size for optimization "~w[o])) +
  ylab("Type II error rate") +
  facet_wrap(~method, labeller="label_both")

@@

cvcs <- cvcs %>% mutate(
  method.wo = paste(method, wo, sep= " / "),
  method.wo = factor(method.wo, 
                     levels = c("CS1 / 90", "CS1 / 120", "CS1 / 150", "CS1 / 180", "CS1 / 210", "CS2 / 90", "CS2 / 120", "CS2 / 150", "CS2 / 180", "CS2 / 210"))
)
cvcs %>% filter(!wo %in% c(130, 140) ) %>%
  ggplot(aes(d = as.numeric(match), m = 1-p_value)) + 
  geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6, 
           aes(colour = method.wo)) +
  facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) +
  scale_alpha_discrete(range=c(0.1, 0.6)) +
  scale_colour_manual(expression("Method /\nOptimization\nwindow size"~w[o]), values=cols) +
  theme_bw() +
  xlab("False Positives Rate (1 - Specificity)") +
  ylab("True Positives Rate (Sensitivity)") +
  guides(
   colour = guide_legend(ncol=2)
  ) +
  geom_abline(slope=-1, intercept = 1, linetype = 3, 
              colour="grey30", size = 0.25)

@@

>>>>

errorsCV2 <- rbind(errorrate(cvcs2, 0.001), 
                errorrate(cvcs2, 0.005), 
                errorrate(cvcs2, 0.01), 
                errorrate(cvcs2, 0.05))
errorsCV2$method <- "CS2"
errorsCV1 <- rbind(errorrate(cvcs1, 0.001), 
                errorrate(cvcs1, 0.005), 
                errorrate(cvcs1, 0.01), 
                errorrate(cvcs1, 0.05))
errorsCV1$method <- "CS1"

errorsCV <- rbind(errorsCV1, errorsCV2)
errorsCV %>% 
  ggplot(aes(x = wo, y = beta, colour=factor(alpha))) + 
  geom_point(aes(shape=factor(wv)), size=2.5) +
  geom_smooth(aes(group=alpha), se=FALSE, method="loess", span=1,  size=.7) +
  theme_bw() +
  scale_colour_brewer(expression("Nominal\ntype I error"~alpha),
                      palette="Set2") +
  scale_shape_discrete(expression("Size of\nvalidation window "~w[v])) +
  xlab(expression("Window size for optimization "~w[o])) +
  ylab("Type II error rate") +
  facet_wrap(~method, labeller="label_both")
```

@@

cvcs <- cvcs %>% mutate(
  method.wo = paste(method, wo, sep= " / "),
  method.wo = factor(method.wo, 
                     levels = c("CS1 / 90", "CS1 / 120", "CS1 / 150", "CS1 / 180", "CS1 / 210", "CS2 / 90", "CS2 / 120", "CS2 / 150", "CS2 / 180", "CS2 / 210"))
)

cvcs %>% filter(!wo %in% c(130, 140) ) %>%
  ggplot(aes(d = as.numeric(match), m = 1-p_value)) + 
  geom_roc(labels=FALSE, n.cuts=100, pointsize=0, size=0.6, 
           aes(colour = method.wo)) +
  facet_wrap(~wv, labeller = label_bquote(w[v] == .(wv))) +
  scale_alpha_discrete(range=c(0.1, 0.6)) +
  scale_colour_manual(expression("Method /\nOptimization\nwindow size"~w[o]), values=cols) +
  theme_bw() +
  xlab("False Positives Rate (1 - Specificity)") +
  ylab("True Positives Rate (Sensitivity)") +
  guides(
   colour = guide_legend(ncol=2)
  ) +
  geom_abline(slope=-1, intercept = 1, linetype = 3, 
              colour="grey30", size = 0.25)

@@

```{r eer, fig.height = 3.25, fig.width = 8, out.width='\\textwidth', fig.cap="Comparison of results for CS1 and CS2 using area under the curve (AUC) and equal error rates (EER). "}
auc <- function(tpr, fpr){
  # from http://blog.revolutionanalytics.com/2016/11/calculating-auc.html
  # inputs already sorted, best scores first 
  dfpr <- c(diff(fpr), 0)
  dtpr <- c(diff(tpr), 0)
  sum(tpr * dfpr) + sum(dtpr * dfpr)/2
}

eer <- function(tpr, fpr) {
  diff = abs(fpr-1+tpr)
  idx <- which.min(diff)
  EER <- 0.5*(fpr[idx]+1-tpr[idx])
  EER
}

EER <- function(score, match) {
  bstats <-data.frame(score, match)
  bstats <- bstats %>% arrange(desc(score))

  bstats <- bstats %>% mutate(
    tpr=cumsum(match)/sum(match), 
    fpr=cumsum(!match)/sum(!match)
  ) %>% unique()

  with(bstats, eer(tpr, fpr))
}

AUC <- function(score, match) {
  bstats <-data.frame(score, match)
  bstats <- bstats %>% arrange(desc(score))

  bstats <- bstats %>% mutate(
    tpr=cumsum(match)/sum(match), 
    fpr=cumsum(!match)/sum(!match)
  ) %>% unique()

  with(bstats, auc(tpr, fpr))
}

roc.summary <- cvcs %>% group_by(wo, wv, method) %>%
  summarize(
    eer = EER(1-p_value, match),
    auc = AUC(1-p_value, match)
  )
p1 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=eer, shape=factor(wv))) +
  geom_point() +
  facet_wrap(~method, labeller="label_both") + 
  theme_bw() +
  scale_shape_discrete("Size of \nvalidation window") +
  ylab("Equal Error rate (EER)") +
  xlab("Size of optimization window in pixels") +
  theme(legend.position = c(0.3, 0.2))

p2 <- roc.summary %>% filter(!(wo %in% c(130, 140))) %>%
  ggplot(aes( x = wo, y=auc, shape=factor(wv))) +
  geom_point() +
  facet_wrap(~method, labeller="label_both") + 
  theme_bw() +
  scale_shape_discrete("Size of \nvalidation window") +
  ylab("Area under the curve (AUC)") +
  xlab("Size of optimization window in pixels")+
  theme(legend.position = c(0.7, 0.2))
grid.arrange(p1, p2, nrow=1)
```
